Decompositions of congruence subgroups of Chevalley groups

### 2015-10-09 stability-conditions.tex
Rewritten example 3.6

Let $R$ be a commutative ring. We denote by $\Max(R)$ its \emph{maximum spectrum} i.e. the set of maximal ideals of $R$ equipped with Zariski topology. For a topological space $X$ denote by $\dim(X)$ its usual topological dimension.
%The following result gives an upper estimate for the absolute stable rank (see~).
%\begin{prop}\label{prop:asr-dimmax}
%Let $R$ be a commutative ring whose maximum spectrum $\Max(R)$ can be covered with a finite family of subsets 
%$X_1\ldots, X_m$ such that $\dim(X_i)\leqslant d$ for every $i$. Then $\asr(A)\leqslant d+1$.
%\end{prop}
From the definition of $\asr(I)$ and \cite[Theorem~2.3]{EO} (or~\cite[Theorem~3.7]{MKV}) it follows that
\begin{equation} \label{sr-estimates} \sr(I)\leqslant\asr(I)\leqslant\asr(R)\leqslant \dim(\Max(R))+1\leqslant\dim(\Spec(R))+1. \end{equation}
There exist examples of rings for which the first inequality in the above formula is strict.
More precisely, there are rings of stable rank $1$ but of absolute stable rank $\geqslant2$, see~\cite{MKV}.
On the other hand, \cite[Theorem~1.3]{MKV} shows that $\sr(R)=\asr(R)$ for any principal ideal ring $R$.
\begin{example}
\eqref{sr-estimates} shows in particular, that every Dedekind domain $R$ satisfies $\asr(R)\leqslant2$, and that $\sr(A[x])\leqslant3$ for every locally principal ring $A$. It is shown in \cite{GruMenVasSL2ZxSL2kxy} that $\sr(\mathbb{Z}[x])=3$. Note that $\sr(\mathbb{Z}[x_1,\ldots,x_n])=n+1$ if $n>1$ \cite[Example after Theorem~18.2]{VasSusSerreProblem}.

Another corollary of \eqref{sr-estimates} is an estimate for the stable rank of the polynomial rings over a field.
Namely, $\sr(F[x_1,\ldots,x_n])\leqslant n+1$ for any field $F$.
For certain classes of fields this upper estimate can be further improved.
For example, if $F$ is algebraic over a finite field then $\sr(F[x_1,\ldots,x_n])\leqslant n$ by~\cite[Corollary~17.4]{VasSusSerreProblem}.
On the other hand, it is known that $\sr(F[x,y])=3$ for every $F$ with $\K_2^M(F)\neq0$ \cite{KrusenmeyerThesis} and $\sr(F[x_1,\ldots,x_n])=n+1$ for $F\subseteq\mathbb{R}$, see~\cite[Theorem~8]{Va71}.
\end{example}

>>>>

Let $R$ be a commutative ring. We denote by $\Max(R)$ its \emph{maximum spectrum} i.e. the set of maximal ideals of $R$ equipped with Zariski topology. For a topological space $X$ denote by $\dim(X)$ its usual topological dimension.
%The following result gives an upper estimate for the absolute stable rank (see~).
%\begin{prop}\label{prop:asr-dimmax}
%Let $R$ be a commutative ring whose maximum spectrum $\Max(R)$ can be covered with a finite family of subsets 
%$X_1\ldots, X_m$ such that $\dim(X_i)\leqslant d$ for every $i$. Then $\asr(A)\leqslant d+1$.
%\end{prop}
From the definition of $\asr(I)$ and \cite[Theorem~2.3]{EO} (or~\cite[Theorem~3.7]{MKV}) it follows that
\begin{equation} \label{sr-estimates} \sr(I)\leqslant\asr(I)\leqslant\asr(R)\leqslant \dim(\Max(R))+1\leqslant\dim(\Spec(R))+1. \end{equation}
There exist examples of rings for which the first inequality in the above formula is strict.
More precisely, there are rings of stable rank $1$ but of absolute stable rank $\geqslant2$, see~\cite{MKV}.
On the other hand, \cite[Theorem~1.3]{MKV} shows that $\sr(R)=\asr(R)$ for any principal ideal ring $R$.
\begin{example}
From inequality~\eqref{sr-estimates} one concludes, in particular, that 
\begin{itemize}
 \item $\asr(F[x_1,\ldots,x_n])\leqslant n+1$ for any field $F$;
 \item $\asr(A[x])\leqslant3$ for every locally principal ring $A$;
 \item $\asr(R)\leqslant2$ for every Dedekind domain $R$.
\end{itemize}
Notice also that $\sr(\mathbb{Z}[x])=3$ (see \cite{GruMenVasSL2ZxSL2kxy}) and $\sr(\mathbb{Z}[x_1,\ldots,x_n])=n+1$ for $n>1$ (see example after~\cite[Theorem~18.2]{VasSusSerreProblem}).

For certain classes of fields the estimate for the stable rank of $R = F[x_1,\ldots, x_n]$ can be further improved.
For example, if $F$ is algebraic over a finite field then $\sr(R)\leqslant n$ by~\cite[Corollary~17.4]{VasSusSerreProblem}.
On the other hand, it is known that $\sr(F[x,y])=3$ for every $F$ with $\K_2^M(F)\neq0$ (see~\cite{KrusenmeyerThesis}) and $\sr(R)=n+1$ for $F\subseteq\mathbb{R}$, see~\cite[Theorem~8]{Va71}.
\end{example}
